---
output: 
  html_document:
    toc: yes
    toc_float:
      collapsed: false
pagetitle: final_project_report
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```

# Information on data:
The following data is on New Orleans tornado building damage during December 2022. This data was obtained from Verisk Analytics and it was derived computer vision and machine learning using post-catastrophe aerial imagry data. There are approximately 42,000 buildings in this dataset. 

---

```{r, include = FALSE}
knitr::opts_chunk$set(echo = FALSE,warning = FALSE,message = FALSE)
library(tidyverse)
library(janitor)
library(leaflet)
library(easystats)
library(modelr)
library(ggmap)
library(sf)
library(GGally)
library(plotly)
library(kableExtra)
library(patchwork)
theme_set(theme_minimal())
```

# Clean data:

I converted roof_solar into a T/F statement, by converting "SOLAR PANEL" to TRUE and "NO SOLAR PANEL" to FALSE. In addition to this, I converted the roof shapes that the computer wasn't very sure about (up to a 20% chance of being incorrect) into NA. There were some cells in damage_level where they were filled with an empty character, so I converted that into NA as well. I then separated longitude and latitude so that it could be easily read into leaflet. 

```{r}
df <- read.csv("clean_data.csv") %>% 
  janitor::clean_names() %>% 
  mutate(roofsolar = case_when(roofsolar == "SOLAR PANEL" ~ TRUE)) %>% # This column was just "yes" or "no"
  mutate(roofshape = ifelse(roofshascr < 0.80, NA, roofshape)) %>% # I only want to include roofshapes that the computer was highly confident about
  select(-c(roofshascr, roofcondit_discolordetect, roofcondit_discolorscore, roofcondit_discolorpercen, trampscr, roofcondit_tarppercen))

df$rooftopgeo <- gsub("POINT \\(|\\)", "", df$rooftopgeo) # clear unnecessary words to obtain long and lat

df <- df %>% # Lat -> y-axis; Long -> x-axis
  separate(rooftopgeo, into = c("long", "lat"), sep = " ", convert = TRUE)

df$damage_level <- ifelse(df$damage_level == "", NA, df$damage_level) # Turn blank observations into NA
```

```{r}
# df %>%
#   kable() %>%
#   kable_classic(lightable_options = "hover") %>%
#   scroll_box(height = "200px")
```


# Define damage categories: 

Separate points based on damage. The majority of the points (3852) exhibited a catastrophe score of 0.

```{r}
# Clear the empty damage ratings: I want to map this 
# There are more data points that exhibit no damage
mostdamage <- df %>% filter(catastrophescore >= 50)
middamage <- df %>% filter(catastrophescore < 50 & catastrophescore > 25)
leastdamage <- df %>% filter(catastrophescore < 25 & catastrophescore != 0)
nodamage <- df %>% filter(catastrophescore == 0)
decimated <-df %>% filter(catastrophescore == 100)

# # Might work better based on quartiles of attributes selected
# mostdamage <- df %>% filter(CATASTROPHESCORE >= 50)
# middamage <- df %>% filter(CATASTROPHESCORE < 50 & CATASTROPHESCORE >= 15)
# lessdamage <- df %>% filter(CATASTROPHESCORE < 15 & CATASTROPHESCORE >= 2)
# minimaldamage <- df %>% filter(CATASTROPHESCORE == 1)

```

# Show maps (I want a good distribution of damage and no damage, I'll have to see if I want to base it on different factors)

See if there are better markers to use for this: see If i can add labels with the long, lat, roof shape, catastrophe_score,
```{r}
alldamage <- leaflet() %>%
  addTiles() %>%  # Add default OpenStreetMap map tiles
  addCircleMarkers(lng=mostdamage$long, lat=mostdamage$lat, color = "red", radius = 2) %>%
  addCircleMarkers(lng=middamage$long, lat=middamage$lat, color = "orange", radius = 2) %>%
  addCircleMarkers(lng=leastdamage$long, lat=leastdamage$lat, color = "blue", radius = 2) %>%
  addCircleMarkers(lng=nodamage$long, lat=nodamage$lat, color = "gray", radius = .5)
alldamage  # Print the map
```
# These are the buildings that sustained damage

```{r}
damage <- leaflet() %>%
  addTiles() %>%  # Add default OpenStreetMap map tiles
  addCircleMarkers(lng=mostdamage$long, lat=mostdamage$lat, color = "red", radius = 2) %>% 
  addCircleMarkers(lng=middamage$long, lat=middamage$lat, color = "orange", radius = 2) %>% 
  addCircleMarkers(lng=leastdamage$long, lat=leastdamage$lat, color = "blue", radius = 2)
damage  # Print the map

damage <- leaflet() %>%
  addTiles() %>%  # Add default OpenStreetMap map tiles
  addMarkers(lng=mostdamage$long, lat=mostdamage$lat) %>%
  addMarkers(lng=leastdamage$long, lat=leastdamage$lat)
damage  # Print the map
```
destryed (i don't think this is working)
```{r}
# ONes that were completely destroyed
wow <- leaflet() %>%
  addTiles() %>%  
  addCircleMarkers(lng=decimated$long, lat=decimated$lat, color = "red", radius = 2) %>% 
  addLabelOnlyMarkers(data = decimated,
                      label = ~paste("Roof shape:", roofshape, "<br>",
                                     "Roof material:", roofmateri, "<br>",
                                     "Catastrophe score:", catastrophescore, "<br>",
                                     "Latitude:", lat, "<br>",
                                     "Longitude:", long),
                      labelOptions = labelOptions(noHide = TRUE, textOnly = FALSE, direction = "auto"))

```



```{r}
leastdamage <- leaflet() %>%
  addTiles() %>%  # Add default OpenStreetMap map tiles
  addMarkers(lng=leastdamage$long, lat=leastdamage$lat)
leastdamage

```

```



leaflet() %>% 
  addProviderTiles("Stamen.Toner") %>% 
  addPolygons(data = shape, fill = NA, color = "grey") %>% 
  addCircleMarkers(data = points, color = "red", radius = 3)


df$catastrophescore %>% unique()
```{r}
# This is a reduced version of the dataset for faster loading times
practice <- df[1:20,]
m <- leaflet(data = practice) %>%
  addTiles() %>%  # Add default OpenStreetMap map tiles
  addMarkers(lng=practice$long, lat=practice$lat, popup="Example")
m  # Print the map

```

Introduce best models




INtroduce predictions


Interpretations of predictions and models used :) amazing owowowowowoow
